TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2018-09-20 paper/draft.tex
describe likelihood fn

Following from the procedure of \citealt{Ness:2015} and \citealt{Ho:2017a} we adopt a simple linear model that assumes the flux at each pixel of the spectrum can be parameterized as a function of a label vector $\ell$ and and coefficient vector \emph{$\theta$}. So for each star \emph{n}, at pixel wavelength \emph{$\lambda$}, we state that the measured flux for a star at a given pixel is the sum of the coefficient and label product, and observational noise:
\begin{equation}
	f_{n\lambda} = \theta_{\lambda}^{T} \cdot \ell_{n} + {\rm noise}
\end{equation}
We use the noise model is ${\rm noise}=[s_{\lambda}^2 + \sigma_{n\lambda}^2]\xi_{n\lambda}$, where the bracketed term is the root mean sum of the intrinsic scatter of the model at each pixel \emph{$s_{\lambda}$}, and the uncertainty due to instrumental effects \emph{$\sigma_{n\lambda}$}, which is then multiplied times a Gaussian random number $\xi_{n\lambda} \sim \mathcal{N} (0,1)$.

@@

The \emph{training step} consists of optimizing for the coefficient vector and scatter ($\theta_{\lambda}$ and $s_{\lambda}$) given fixed label vector (\emph{$\ell_n$}), and the \emph{test step} consists of optimizing for the labels at fixed $\theta_{\lambda}$ and $s_{\lambda}$.

% JB: The training step description: OPTIMIZE THETA AT FIXED LABELS.
% JB: Test step description: OPTIMIZE LABELS AT FIXED THETA.

Thus \thecannon\ is both a generative model for the stellar spectra, and a method for obtaining
labels from spectra. In the training step, the regression is designed to predict spectral pixels
given labels, by learning zeroth, first, and second derivatives of the data with respect to
the labels. In the test step, the regression is designed to predict labels given the spectral
derivatives.

>>>>

Following from the procedure of \citealt{Ness:2015} and \citealt{Ho:2017a} we adopt a simple linear model that assumes the flux at each pixel of the spectrum can be parameterized as a function of a label vector $\ell$ and and coefficient vector \emph{$\theta$}. So for each star \emph{n}, at wavelength pixel \emph{$\lambda$}, we state that the measured flux for a star at a given pixel is the sum of the coefficient and label product, and observational noise:
\begin{equation}
	f_{n\lambda} = \theta_{\lambda}^{T} \cdot \ell_{n} + {\rm noise} \label{eqn:cannon}
\end{equation} 
We use the noise model is ${\rm noise}=[s_{\lambda}^2 + \sigma_{n\lambda}^2]\xi_{n\lambda}$, where the bracketed term is the root mean sum of the intrinsic scatter of the model at each pixel \emph{$s_{\lambda}$}, and the uncertainty due to instrumental effects \emph{$\sigma_{n\lambda}$}, which is then multiplied times a Gaussian random number $\xi_{n\lambda} \sim \mathcal{N} (0,1)$. Equation \ref{eqn:cannon} corresponds to the single-pixel log-likelihood function:
\begin{equation}
	\ln p(f_{n\lambda}|\theta^T_{\lambda}, \ell_n, s^2_n) = 
	-\frac{1}{2}\frac{[f_{n\lambda} - \theta_{\lambda}^{T} \cdot \ell_{n}]^2}{s_{\lambda}^2 + \sigma_{n\lambda}^2} - \frac{1}{2} \ln(s_{\lambda}^2 + \sigma_{n\lambda}^2) 
	\label{eqn:likelihood}
\end{equation}
the probability density function of the measured flux, given the labels, coefficients and scatters.

@@

The \emph{training step} consists of optimizing the likelihood function (Equation \ref{eqn:likelihood}) for the coefficient vector and scatter ($\theta_{\lambda}$ and $s_{\lambda}$) given fixed label vector (\emph{$\ell_n$}) constructed from the reference labels, and the \emph{test step} consists of optimizing the likelihood function for the labels at fixed $\theta_{\lambda}$ and $s_{\lambda}$ obtained in the training step (see \citealt{Ness:2015} for further description). 
Thus \thecannon\ is both a generative model for the stellar spectra, and a method for obtaining 
labels from spectra. In the training step, the regression is designed to predict spectral pixels
given labels, by learning zeroth, first, and second derivatives of the data with respect to
the labels. In the test step, the regression is designed to predict labels given the spectral
derivatives.
